Improving cropland maps through tight integration of human and machine intelligence

### 2021-02-11 inst/manuscript/figures.Rmd
Function for outputting accuracy reports, scripts for processing accuracy reports, figure for accuracy tables with AOI zones

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(raster)
library(sf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(RColorBrewer)
library(activemapper)
pth <- function(x) file.path(here("inst/manuscript"), x)
```

@@

degrees <- st_read(
  system.file("extdata/degrees.geojson", package = "activemapper")
)
tiles <- st_read(
  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
) 
aois <- st_read(
  system.file("extdata/aois.geojson", package = "activemapper")
)
# ggplot(data = tiles) + 
#   geom_sf(aes(fill = as.factor(aoi1)), col = "transparent")

@@

### Figure 7: Categorical map accuracy

### Figure 8: Cropland maps
```{r, eval=FALSE}
library(stars)
library(ggplot2)
library(aws.s3)
library(RStoolbox)
library(cowplot)

# segment files
segment_files <- dir(here("external/data/results/segments/merged"),
                     pattern = "boundarymerge", full.names = TRUE)
iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file))

@@

# entire cropland fraction
cropfrac <- stars::read_stars(
  system.file("extdata/probs_005.tif", package = "activemapper")
)

@@

# img <- read_stars(impath)
# mxval <- st_apply(img, 3, max) %>% pull(max)
# imgngb <- st_rgb(img[, , , 4:2], 3, maxColorValue = mxval[4:2])
vpal <- viridis::viridis(20)
inset_cols <- c("red", "orange", "pink", "brown")
fld_tiles <- list(flds_tile1, flds_tile2, flds_tile3, flds_tile4)
isets <- lapply(1:4, function(x) {  # x <- 1
  imgr <- brick(impath[x])[[2:4]]
  probr <- stars::read_stars(prob_tiles$impath[x])
  xylims <- st_bbox(imgr)
  iset <- ggplot() + 
    ggRGB(imgr, maxpixels = 50000, stretch = "lin", ggLayer = TRUE) +
    geom_sf(data = fld_tiles[[x]], fill = "transparent", color = "yellow", 
            size = 0.2) + 
    geom_sf(data = tile_ss[x, ], color = inset_cols[x], fill = "transparent") + 
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void()
  iset2 <- ggplot() + 
    geom_stars(data = probr) +
    scale_fill_gradientn(name = "", colours = vpal, limits = c(0, 100)) +
    geom_sf(data = tile_ss[x, ],  color = inset_cols[x], fill = "transparent") +
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void() + theme(legend.position = "none")
  cowplot::plot_grid(iset2, iset, ncol = 1, align = "hv")
})
# isets[[2]]

@@

# cropland fractions
p <- ggplot(africa) + 
  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
  geom_stars(data = cropfrac) +
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
  geom_sf(data = tile_ss, fill = inset_cols, col = inset_cols, size = 0.1) +
  # geom_sf(data = tile_ss, fill = "red", col = "red") +
  scale_fill_viridis_c(name = "% cover", na.value = "transparent") +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  theme_void() + 
  theme(legend.position = c(0.65, 0.07), legend.direction = "horizontal", 
        legend.key.height = unit(0.3, "cm")) 
        # legend.text = element_text(colour = "white"), 
        # legend.title = element_text(colour = "white"))
# with insets added
wh <- c(0.4, 0.4)
lx <- -0.1
rx <- 0.7
inset_map <- ggdraw() +
  draw_plot(p) +
  draw_plot(isets[[1]], x = lx, y = 0.6, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[2]], x = rx, y = 0.51, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[3]], x = lx, y = 0.17, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[4]], x = rx, y = 0.07, width = wh[1], height = wh[2])
# inset_map

@@

  # geom_stars(data = imgngb) + scale_fill_identity()
  # geom_sf(data = flds_tile)

plot(flds_tile[1])

fboundary <- read_sf(segment_files$file[1])

pal <- RColorBrewer::brewer.pal(n = 9, "Spectral")[c(1, 2, 3, 4, 9)]

# Panels, organized top to bottom
# Panel 1: Cropland probabilities, with 3 m probability side maps (tiles)
p <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +  
  geom_stars(data = fdensity * 100) + 
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") + 
  scale_fill_viridis_c(name = "% cover", na.value = "transparent") + 
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  theme_void()

# Panel 2: Cropland cover, with polygon side maps over composites (tiles)
p <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +  
  geom_stars(data = fdensity * 100) + 
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") + 
  geom_sf(data = tile_ss, fill = "red", col = "red") + 
  scale_fill_viridis_c(name = "% cover", na.value = "transparent") + 
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  theme_void()
p

# Panel 3: Average field size (kriging adjusted?),

# Panel 4: Field number (kriging adjusted)

>>>>

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(activemapper)
library(raster)
library(sf)
library(stars)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(here)
library(glue)
library(aws.s3)
library(RStoolbox)
# library(gridExtra)
library(ggpmisc)
library(patchwork)

pth <- function(x) file.path(here("inst/manuscript"), x)
```

@@

degrees <- st_read(
  system.file("extdata/degrees.geojson", package = "activemapper")
)
tiles <- st_read(
  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
) 
aois <- st_read(
  system.file("extdata/aois.geojson", package = "activemapper")
)
aez <- st_read(
  system.file("extdata/ghana_aez.geojson", package = "activemapper")
)

# ggplot(data = tiles) + 
#   geom_sf(aes(fill = as.factor(aoi1)), col = "transparent")

@@

### Figure 7: Categorical map accuracy
Place map accuracy tables and zone report into single figure, for AOI zones.

```{r, eval=FALSE}
data("accuracy_reports")

# zone maps
zone_grps <- list("aoi" = list(1:3, 4:9, c(10:11, 13, 14, 16), c(12, 15)),
                  "aez" = list(1, 2:4, 5, 6:7))
aoi_zones <- aois %>% 
  mutate(zone = case_when(
    aois %in% zone_grps$aoi[[1]] ~ "Zone 1", 
    aois %in% zone_grps$aoi[[2]] ~ "Zone 2", 
    aois %in% zone_grps$aoi[[3]] ~ "Zone 3", 
    aois %in% zone_grps$aoi[[4]] ~ "Zone 4")
  ) %>%
  group_by(zone) %>% 
  summarize(area = sum(area))

# RF accuracy
rfreports <-  lapply(accuracy_reports$zones$aoi, function(x) x$rf)
rfreports[[length(rfreports) + 1]] <- accuracy_reports$ghana$rf
rf_tabs <- lapply(1:5, function(i) {  # i <- 2
  m <- rfreports[[i]]
  d <- data.frame(m, check.names = FALSE)
  rbind(d, rep("", ncol(d)))
}) %>% do.call(rbind, .)
zvec <- rep("", nrow(rf_tabs))
znames <- paste(c(paste("Zone", 1:4), "Ghana"), "  ")
zvec[c(3, seq(8, length(zvec), by = 5))] <- znames
rf_tabs <- cbind(zvec, rf_tabs)
colnames(rf_tabs) <- c("", colnames(rfreports$North))

rf_tbl <- tibble(x = 1, y = 10, dat = list(rf_tabs))

# segmentation accuracy
segreports <-  lapply(accuracy_reports$zones$aoi, function(x) x$seg)
segreports[[length(segreports) + 1]] <- accuracy_reports$ghana$seg
seg_tabs <- lapply(1:5, function(x) {  # x <- 1
  m <- segreports[[x]]
  d <- data.frame(m, check.names = FALSE)
  rbind(d, rep("", ncol(d)))
}) %>% do.call(rbind, .)
colnames(seg_tabs) <- colnames(rfreports$North)
seg_tbl <- tibble(x = 1, y = 10, dat = list(seg_tabs))

# Assemble plots with maps
# Table theme
tt <- ttheme_gtminimal(
  base_size = 4.5, 
  padding = unit(c(0.2, 0.2), "char"),
  core = list(fg_params = list(hjust = 1, x = 1))
)

m <- ggplot(aoi_zones) + geom_sf(lwd = 0.1) + 
  geom_sf_text(data = st_centroid(aoi_zones), aes(label = zone), size = 2) + 
  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

t1 <- ggplot() + 
  geom_table(data = rf_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
  ggtitle("Per-pixel classification") + 
  theme_void() + 
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 6))
t2 <- ggplot() + 
  geom_table(data = seg_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
  ggtitle("Segmentation") +
  theme_void() + 
  theme(plot.margin = unit(c(0,0,0,0), "cm"), 
        plot.title = element_text(hjust = 0.5, vjust = 1, size = 6))

# gp <- cowplot::plot_grid(t1, m, t2, nrow = 1, align = "hv", 
#                          # rel_widths = c(0.2, 0.4, 0.5, 0.4, 0.2),
#                          rel_widths = c(1, 1, 0.8),
#                          # rel_heights = c(0.5, 1.5, 0.5),
#                          scale = c(0.25, 0.95, 0.25),
#                          label_size = 5, 
#                          label_x = 0.25, 
#                          labels = c("Per-pixel classification", "",
#                                     "Segmentation"))
gp <- t1 + m + t2
ggsave(gp, filename = pth("figures/figure7.png"), width = 7.25, height = 2.35, 
       units = "in", dpi = 300)
```

### Figure 8: Cropland maps
```{r, eval=FALSE}
# segment files
segment_files <- dir(here("external/data/results/segments/merged"),
                     pattern = "boundarymerge", full.names = TRUE)
iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file))

@@

# load in Ghana cropland fractions, mask out tiles over ocean, convert to stars
# percent
cropfrac <- raster(
  system.file("extdata/cropfrac_005.tif", package = "activemapper")
)
cropfrac <- africa %>% 
  filter(name %in% c("Ghana", "Togo", "Cte d'Ivoire", "Burkina Faso")) %>% 
  mask(cropfrac, .)
cropfrac <- st_as_stars(cropfrac * 100)

@@

# process inset maps
# img <- read_stars(impath)
# mxval <- st_apply(img, 3, max) %>% pull(max)
# imgngb <- st_rgb(img[, , , 4:2], 3, maxColorValue = mxval[4:2])
vpal <- viridis::viridis(20)
inset_cols <- c("red", "orange", "pink", "brown")
fld_tiles <- list(flds_tile1, flds_tile2, flds_tile3, flds_tile4)
isets <- lapply(1:4, function(x) {  # x <- 1
  imgr <- brick(impath[x])[[2:4]]
  probr <- stars::read_stars(prob_tiles$impath[x])
  xylims <- st_bbox(imgr)
  iset <- ggplot() + 
    ggRGB(imgr, maxpixels = 50000, stretch = "lin", ggLayer = TRUE) +
    geom_sf(data = fld_tiles[[x]], fill = "transparent", color = "yellow", 
            size = 0.2) + 
    geom_sf(data = tile_ss[x, ], color = inset_cols[x], fill = "transparent") + 
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void()
  iset2 <- ggplot() + 
    geom_stars(data = probr) +
    scale_fill_gradientn(name = "", colours = vpal, limits = c(0, 100)) +
    geom_sf(data = tile_ss[x, ],  color = inset_cols[x], fill = "transparent") +
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void() + theme(legend.position = "none")
  cowplot::plot_grid(iset2, iset, ncol = 1, align = "hv")
})
# isets[[2]]

@@

# cropland fractions
p <- ggplot(africa) + 
  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
  geom_stars(data = cropfrac) +
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
  geom_sf(data = tile_ss, fill = inset_cols, col = inset_cols, size = 0.1) +
  # geom_sf(data = tile_ss, fill = "red", col = "red") +
  scale_fill_viridis_c(name = "%", na.value = "transparent") +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  theme_void() + 
  theme(legend.position = c(0.64, 0.08), legend.direction = "horizontal", 
        legend.key.height = unit(0.3, "cm"), 
        legend.text = element_text(colour = "black"),
        legend.title = element_text(size = 10, colour = "black"))
# with insets added
wh <- c(0.4, 0.4)
lx <- -0.1
rx <- 0.7
inset_map <- ggdraw() +
  draw_plot(p) +
  draw_plot(isets[[1]], x = lx, y = 0.6, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[2]], x = rx, y = 0.51, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[3]], x = lx, y = 0.17, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[4]], x = rx, y = 0.07, width = wh[1], height = wh[2])
# inset_map

@@

### 2021-02-11 inst/manuscript/manuscript.Rmd
Function for outputting accuracy reports, scripts for processing accuracy reports, figure for accuracy tables with AOI zones

>>>>

```{r mainmap, echo = FALSE, out.width="90%", fig.cap="The distribution of arable croplands in Ghana. The main map shows the percentage of croplands in each 0.005 degree grid cell, derived from the predicted cropland probabilities. The insets  on the margins illustrate predicted probabilities (top map in each couplet) at original resolution (0.000025 degrees) and segmented field boundaries overlaid on the dry season PlanetScope composite, for four separate tiles. Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure8.png')
```
